---
title: "gse37614 batch correction"
author: "Clodagh Murray"
date: "5/19/2022"
output: html_document
---


### Libraries
```{r, message =F}
library(Biobase)
library(ggpubr)
library(data.table)
library(DT)
library(dplyr)
library(ggpubr)
library(ggplot2)
library(limma)
```

## Raw Data Staging
```{r}
SDRF <- read.delim("/home/clodagh/MASTERS PROJECT/GSE37614/E-GEOD-37614.sdrf.txt")
rownames(SDRF) <- SDRF$Array.Data.File
DT::datatable(SDRF, options = list(scrollX = TRUE, pageLength = 19, scroller = TRUE))
```

## Signal Files
```{r}
#function to stick files together in a dataframe
gen_raw <- function(path, pattern){
  files = list.files(path, pattern, full.names=T, recursive=T, include.dirs=T)
  mat = as.data.frame(do.call(cbind, lapply(files, function(x) data.table::fread(x, stringsAsFactors=FALSE))))
  ID_REF = mat[,1]
  mat = as.data.frame(mat[,!grepl("ID_REF", colnames(mat))])
  rownames(mat) = ID_REF
  return(mat)
}


raw.dat <- gen_raw("/home/clodagh/MASTERS PROJECT/GSE37614/raw/", "\\.txt$")
dim(raw.dat)
head(raw.dat)
```

```{r}
# append correct sample names to each column
file_names <- list.files("/home/clodagh/MASTERS PROJECT/GSE37614/raw/", "\\.txt$", full.names=T, recursive=T, include.dirs=T)
file_names <- sub(".*/", "", file_names)
head(file_names)
```

```{r}
sample_names <- sub("\\_.*", "", file_names)
head(sample_names)
```

```{r}
# There are more columns in raw.dat (23) than there are sample names in the annotation file (19).
length(sample_names) == ncol(raw.dat) #FALSE
```
### 4 samples were analysed in both batches 1 and 2, creating 4 extra samples in the raw data - update metadata to match raw data 

```{r}
SDRF <- read.delim("/home/clodagh/MASTERS PROJECT/GSE37614/E-GEOD-37614.sdrf_fixed.txt")
rownames(SDRF) <- SDRF$Array.Data.File
DT::datatable(SDRF, options = list(scrollX = TRUE, pageLength = 23, scroller = TRUE))
```

## Count matrix matching metadata?
```{r}
## same as first attempt tbh
gen_raw <- function(path, pattern){
  files = list.files(path, pattern, full.names=T, recursive=T, include.dirs=T)
  mat = as.data.frame(do.call(cbind, lapply(files, function(x) data.table::fread(x, stringsAsFactors=FALSE))))
  ID_REF = mat[,1]
  mat = as.data.frame(mat[,!grepl("ID_REF", colnames(mat))])
  rownames(mat) = ID_REF
  return(mat)
}

raw.dat <- gen_raw("/home/clodagh/MASTERS PROJECT/GSE37614/raw/", "\\.txt$")

## attach sample names to cols
file_names <- list.files("/home/clodagh/MASTERS PROJECT/GSE37614/raw/", "\\.txt$", full.names=T, recursive=T, include.dirs=T)
file_names <- sub(".*/", "", file_names)
sample_names <- sub("\\_.*", "", file_names)
length(sample_names) == ncol(raw.dat)
```

## Append IDs to matrix

```{r}
# we need to append batch ID's to the samplenames - sorry but has to happen for unique colnames. 
SDRF$samp_names <- paste(SDRF$Assay.Name, SDRF$Characteristics..batch., sep=":")
SDRF$samp_names

```
```{r}
# below is most important part to make sure columns match file contents
SDRF <- SDRF[match(file_names, SDRF$Array.Data.File),]
colnames(raw.dat) <- SDRF$samp_names

## the below is correct, check it against the head -n 4 *.txt file command above
head(raw.dat[1:5, 1:5])
```

## Expression Set 

```{r}
SDRF <- Biobase::AnnotatedDataFrame(SDRF)
rownames(SDRF) <- colnames(raw.dat)
my_set <- Biobase::ExpressionSet(assayData = as.matrix(raw.dat), phenoData = SDRF)
my_set
```

## Raw Data Boxplot
```{r}
SDRF@data$Characteristics..class.[23] <- "Her2+"
sml <- as.character(as.numeric(as.factor(SDRF@data$Characteristics..class.)))
gs <- factor(sml)
groups <- SDRF@data$Characteristics..class. # HER2+ER+ sample? 
stupid_orientation <- unique(groups)
stupid_orientation <- c(stupid_orientation[2], stupid_orientation[3], stupid_orientation[1])
levels(gs) <- stupid_orientation
my_set@phenoData$group <- gs
ord <- order(gs)

palette <- palette(c("#1B9E77", "#7570B3", "pink"))

par(mar=c(7,4,2,1))
title <- "GSE37614 Log2 Raw Signals"
boxplot(log2(exprs(my_set[,ord])), boxwex=0.6, notch=T, main=title, outline=FALSE, las=2, col=gs[ord])
legend("topleft", stupid_orientation, fill=palette(), bty="n")
```
## Qunatile normalization

```{r}
quantile_normalisation <- function(df){
  df_rank <- apply(df,2,rank,ties.method="min")
  df_sorted <- data.frame(apply(df, 2, sort))
  df_mean <- apply(df_sorted, 1, mean)

  index_to_mean <- function(my_index, my_mean){
    return(my_mean[my_index])
  }
df_final <- apply(df_rank, 2, index_to_mean, my_mean=df_mean)
  rownames(df_final) <- rownames(df)
  return(df_final)
}

qnorm <- quantile_normalisation(raw.dat)
```

## Boxplot of Qnorm

```{r}
par(mar=c(7,4,2,1))
title <- "GSE37614 Log2 Quantile Norm McDonald"
boxplot(log2(qnorm[,ord]), boxwex=0.6, notch=T, main=title, outline=FALSE, las=2, col=gs[ord])
legend("topleft", stupid_orientation, fill=palette(), bty="n")
```
## Update ExpressionSet Object 

```{r}
exprs(my_set) <- qnorm

## check that actually saved
par(mar=c(7,4,2,1))
title <- "GSE37614 Log2 Quantile Norm McDonald"
boxplot(log2(exprs(my_set[,ord])), boxwex=0.6, notch=T, main=title, outline=FALSE, las=2, col=gs[ord])
legend("topleft", stupid_orientation, fill=palette(), bty="n")
```
## Expression Distribution

```{r}
ex <- exprs(my_set)
par(mar=c(4,4,2,1))
title <- paste ("Normalized Expression Density", sep ="")
limma::plotDensities(ex, group=gs, main=title, legend ="topright")
```

## Probe Filtering

```{r}
# Probe filtering
medians <- rowMedians(ex)

hist(log2(medians + 1), 150, col = "cornsilk1", freq = FALSE, 
             main = "Histogram of the median intensities", 
             border = "antiquewhite4", xlab = "Median intensities")
```
## Remove Multimap probes

```{r}
probes <- rownames(my_set)

library(illuminaHumanv4.db)
```

```{r}
annotation <- AnnotationDbi::select(illuminaHumanv4.db,
                                    keys = probes,
                                    columns = c("SYMBOL", "GENENAME"),
                                    keytype = "PROBEID")
```

```{r}
## 'select()' returned 1:many mapping between keys and columns
annotation <- subset(annotation, !is.na(SYMBOL))

## resolve multi maps. 

ann_grouped <- group_by(annotation, PROBEID)
ann_sum <- dplyr::summarize(ann_grouped, no_of_matches = n_distinct(SYMBOL))

ann_flt <- filter(ann_sum, no_of_matches > 1)

remove_id <- (probes %in% ann_flt$PROBEID)

table(remove_id)
```
# Too many genes

```{r}
annotations <- read.table("/home/clodagh/MASTERS PROJECT/GSE37614/A-GEOD-10558.adf.txt", header=T, sep="\t")

mart <- biomaRt::useMart(biomart = "ensembl", dataset = "hsapiens_gene_ensembl")
attr <- biomaRt::listAttributes(mart)
genbank_map <- biomaRt::getBM(attributes = c("refseq_mrna", "hgnc_symbol", "gene_biotype"), mart = mart, useCache = FALSE)
genbank_map <- genbank_map[!(is.na(genbank_map$refseq_mrna) | genbank_map$refseq_mrna==""),]
genbank_map <- subset(genbank_map, genbank_map$gene_biotype == "protein_coding")

# pop off the version ID
annotations$genbank <- annotations$Reporter.Database.Entry..genbank.

annotations$genbank <- gsub("\\.\\d+$", "", annotations$genbank)
# merge
master <- merge(genbank_map, annotations, by.x="refseq_mrna", by.y="genbank")
```

```{r}
master_sub <- master[,c(4,2)]
master_sub <- master_sub[!(is.na(master_sub$hgnc_symbol) | master_sub$hgnc_symbol==""),]

# remove multi mapping probes
master_sub <- master_sub %>% distinct(hgnc_symbol, .keep_all = T)
dim(master_sub)
```
```{r}
qnorm <- subset(qnorm, rownames(qnorm) %in% master_sub$Reporter.Name)
dim(qnorm)
```

```{r}
master_sub <- master_sub[match(rownames(qnorm), master_sub$Reporter.Name),]
rownames(qnorm) <- master_sub$hgnc_symbol
head(qnorm[1:3,1:3])
```

```{r}
medians <- rowMedians(qnorm)

hist(log2(medians + 1), 150, col = "cornsilk1", freq = FALSE, 
             main = "Histogram of the median intensities", 
             border = "antiquewhite4", xlab = "Median intensities")
```

```{r}
medians <- log2(rowMedians(qnorm) + 1)
qnorm <- as.data.frame(qnorm)
qnorm$medians <- medians

qnorm <- subset(qnorm, qnorm$medians > 7.9)
qnorm <- as.matrix(qnorm[,1:(ncol(qnorm)-1)])

hist(log2(qnorm + 1), 150, col = "cornsilk1", freq = FALSE, 
             main = "Histogram of the median intensities", 
             border = "antiquewhite4", xlab = "Median intensities")
```
# Exploratory Data Analysis

## PCA Cancer Subtypes
2 groups - largely driven by batch
```{r}

PCA <- prcomp(t(log2(qnorm)), scale = TRUE)
percentVar <- round(100*PCA$sdev^2/sum(PCA$sdev^2),1)
sd_ratio <- sqrt(percentVar[2] / percentVar[1])

dataGG <- data.frame(PC1 = PCA$x[,1], PC2 = PCA$x[,2],
                    Subtype = my_set$group)


ggscatter(dataGG, x="PC1", y="PC2",
                  color = "Subtype", palette = c("dodgerblue4", "darkorange2", "chartreuse", "purple"),
                  title = "PCA plot log-transformed quantile normalized expression data",
                  subtitle = "Cancer Subtypes",
                  xlab = paste0("PC1, VarExp: ", percentVar[1], "%"),
                  ylab = paste0("PC2, VarExp: ", percentVar[2], "%"),
                  ellipse = FALSE, star.plot = FALSE, # HER2+ER+ sample n=1 - star, ellipse breaks
                  ggtheme = theme_bw()) + 
                  theme(legend.position = "right") + 
                  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```
## PCA Batch Effects
```{r}
dataGG <- data.frame(PC1 = PCA$x[,1], PC2 = PCA$x[,2],
                    Batch = my_set$Characteristics..batch.)


ggscatter(dataGG, x="PC1", y="PC2",
                  color = "Batch", palette = c("dodgerblue4", "darkorange2", "chartreuse", "purple"),
                  title = "PCA plot log-transformed quantile normalized expression data",
                  subtitle = "Batch Effects (Batch 1, Batch 2)",
                  xlab = paste0("PC1, VarExp: ", percentVar[1], "%"),
                  ylab = paste0("PC2, VarExp: ", percentVar[2], "%"),
                  ellipse = TRUE, star.plot = TRUE,
                  ggtheme = theme_bw()) + 
                  theme(legend.position = "right") + 
                  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```
## Pearsons Correlation btween PCs and batch and cancer

```{r}
meta <- my_set@phenoData@data
meta <- meta[,c(4, 5)]
colnames(meta) <- c("Batch", "Cancer_Subtype")


pca <- PCAtools::pca(log2(qnorm + 1), metadata = meta)

PCAtools::eigencorplot(pca, components = PCAtools::getComponents(pca, 1:10),
    metavars = c('Batch','Cancer_Subtype'),
    col = c('white', 'cornsilk1', 'gold', 'forestgreen', 'darkgreen'),
    cexCorval = 1.2, fontCorval = 2, posLab = 'all', rotLabX = 45,
    scale = TRUE,
    main = bquote(PC ~ Pearson ~ r^2 ~ clinical ~ correlates),
    plotRsquared = TRUE,
    corFUN = 'pearson',
    corUSE = 'pairwise.complete.obs',
    corMultipleTestCorrection = 'BH',
    signifSymbols = c('****', '***', '**', '*', ''),
    signifCutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1))
```
## Batch Correction
```{r}
groups <-  c("ER", "Her2", "TNBC")
design <- model.matrix(~0 + groups)
colnames(design) <- c("ER", "Her2", "TNBC")
rm_batch <- limma::removeBatchEffect(log2(qnorm + 1), batch = my_set@phenoData@data$Characteristics..batch., design=design)

fit <- limma::lmFit(rm_batch, design)

# Set up contrasts of interest
cts <- paste(groups, c(tail(groups, -1), head(groups, 1)), sep="-")
cont.matrix <- limma::makeContrasts(contrasts=cts, levels=design)

fit2 <- limma::contrasts.fit(fit, cont.matrix)
fit2 <- limma::eBayes(fit2, 0.01)
res <- limma::decideTests(fit2)
summary(res)
```

```{r}
#Create the top table for the comparison (coef) we want
# coef = 1 (ER-Her2)
top <- topTable(fit2, coef=1, adjust="fdr", sort.by="B", number = Inf)
nrow(top)

#Top Table of significant genes only
up_reg <- subset(top, logFC > 0.25 & P.Value < 0.05)
nrow(up_reg)

```

## Heatmap 

```{r}
up_reg_names <- rownames(up_reg)
e <- rm_batch
sig_exprs <- e[rownames(e) %in% up_reg_names,]

sig_exprs <- t(sig_exprs)
sig_exprs <- scale(sig_exprs, center=T, scale = T)
sig_exprs <- t(sig_exprs)

annot_col <- data.frame(row.names = colnames(rm_batch), Group = factor(my_set$group))

cols <- c("red", "blue", "green")
names(cols) <- c("ER+", "Her2+", "TNBC")
annot_colors <- list(Group = cols)

library(viridis)
```


```{r}
pheatmap::pheatmap(sig_exprs, cluster_cols = T, cluster_rows = T, show_rownames = F,
                   show_colnames = T, annotation_col = annot_col,
                   annotation_colors = annot_colors,
                   col = hcl.colors(100, "Purple-Green",rev=F),
                   cutree_cols = 4)
```


```{r}
#Create an excel-readable table with the specified columns for sig gene only
tT <- subset(top, select=c("logFC","adj.P.Val","P.Value","t"))
write.table(tT, "DEGs_ER-Her2.csv", row.names=F, sep="\t")
# Expression levels for ALL genes

top_all <- topTable(fit2, adjust="fdr", sort.by="B", number=Inf)


hist(top_all$adj.P.Val, col = "grey", border = "white", xlab = "P-adj",
     ylab = "Number of genes", main = "P-adj value distribution")
```


```{r}
#Extract the expression values for the DEGs
rm_batch
sig_exprs <- rm_batch[rownames(rm_batch) %in% rownames(tT),]
dim(sig_exprs)
```
```{r}
# summarize test results as "up", "down" or "not expressed"
dT <- decideTests(fit2, adjust.method="fdr", p.value=0.05)
vennDiagram(dT, circle.col=palette())
```

```{r}
# create Q-Q plot for t-statistic
t.good <- which(!is.na(fit2$F)) # filter out bad probes
qqt(fit2$t[t.good], fit2$df.total[t.good], main="Moderated t statistic")
```


## PCA Cancer Subtype (Corrected)

```{r}
PCA <- prcomp(t(log2(rm_batch)), scale = TRUE)
percentVar <- round(100*PCA$sdev^2/sum(PCA$sdev^2),1)
sd_ratio <- sqrt(percentVar[2] / percentVar[1])

dataGG <- data.frame(PC1 = PCA$x[,1], PC2 = PCA$x[,2],
                    Subtype = my_set$group)


ggscatter(dataGG, x="PC1", y="PC2",
                  color = "Subtype", palette = c("dodgerblue4", "darkorange2", "chartreuse", "purple"),
                  title = "PCA plot log-transformed batch corrected data",
                  subtitle = "Cancer Subtypes",
                  xlab = paste0("PC1, VarExp: ", percentVar[1], "%"),
                  ylab = paste0("PC2, VarExp: ", percentVar[2], "%"),
                  ellipse = F, star.plot = F, 
                  ggtheme = theme_bw()) + 
                  theme(legend.position = "right") + 
                  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```
## PCA Batch Effect (Correcter)

```{r}
dataGG <- data.frame(PC1 = PCA$x[,1], PC2 = PCA$x[,2],
                    Batch = my_set$Characteristics..batch.)


ggscatter(dataGG, x="PC1", y="PC2",
                  color = "Batch", palette = c("dodgerblue4", "darkorange2", "chartreuse", "purple"),
                  title = "PCA plot log-transformed batch corrected data",
                  subtitle = "Batch Effects: Batch1, Batch2",
                  xlab = paste0("PC1, VarExp: ", percentVar[1], "%"),
                  ylab = paste0("PC2, VarExp: ", percentVar[2], "%"),
                  ellipse = T, star.plot = T, 
                  ggtheme = theme_bw()) + 
                  theme(legend.position = "right") + 
                  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

##  Volcano Plot
```{r}
ct <- 1
# choose contrast of interest
volcanoplot(fit2, coef=ct, main=colnames(fit2)[ct], pch=20,
            highlight=length(which(dT[,ct]!=0)), names=rep('+', nrow(fit2)))
```





