---
title: "GSE108565"
author: "Clodagh Murray"
date: "6/12/2022"
output: html_document
---

```{r, message = F}
library(GEOquery)
library(limma)
library(dplyr)
library(gplots)
library(Matrix)
library(ggpubr)
library(ggplot2)
library(pheatmap)
library(topGO)
library(Rgraphviz)
library(org.Hs.eg.db)
library(ggplot2)
library(enrichR)
library(enrichplot)
library(DT)
library(biomaRt)
```

# Load in data
Expression data were normalized through quantile normalization, and the Robust Multichip Average (RMA) algorithm was included in the NimbleScan software

```{r}
# load series and platform data from GEO
gset <- getGEO("GSE108565", GSEMatrix =TRUE, AnnotGPL=FALSE)
gset <- gset[[1]]
#write.csv(x = x, file = "/home/clodagh/MASTERS PROJECT/GSE108565.expression.matrix.csv", quote = F, row.names = T) #export expression matrix in file (.csv format).
#gset <- read.csv("/home/clodagh/MASTERS PROJECT/GSE108565.expression.matrix.csv")
meta <- pData(gset)
meta <- meta[ ,c("geo_accession","neo-adjuvant chemotherapy:ch1")]
#write.csv(meta , file = "/home/clodagh/MASTERS PROJECT/GSE37614/GSE108565_metadata.csv", quote = F, row.names = F)
```

## General Data pre-processing 
```{r}
sml <- as.character(as.numeric(as.factor(meta$`neo-adjuvant chemotherapy:ch1`)))
gs <- factor(sml)
groups <- meta$`neo-adjuvant chemotherapy:ch1`
orientation <- unique(groups)
levels(gs) <- orientation
#gset@phenoData$group <- gs
ord <- order(gs)
palette <- palette(c("#1B9E77", "#7570B3"))
par(mar=c(7,4,2,1))
title <- "GSE108565 Log2 Normalized Signals"
boxplot(exprs(gset)[,ord], boxwex=0.6, notch=T, main=title, outline=FALSE, las=2, col=gs[ord])
legend("topleft",orientation, fill=palette(), bty="n")
```

# Plot expression density 
```{r}
ex <- qnorm
par(mar=c(4,4,2,1))
title <- paste ("Normalized Expression Density", sep ="")
limma::plotDensities(ex, group=gs, main=title, legend ="topright")
```



# Probe Filtering
## Remove multi-mapping probes
```{r}
annotations <- read.table("/home/clodagh/MASTERS PROJECT/GSE108565/A-GEOD-16025.adf.txt", header=T, sep="\t")

mart <- biomaRt::useMart(biomart = "ensembl", dataset = "hsapiens_gene_ensembl")
attr <- biomaRt::listAttributes(mart)
genbank_map <- biomaRt::getBM(attributes = c("refseq_mrna", "hgnc_symbol", "gene_biotype"), mart = mart, useCache = FALSE)
genbank_map <- genbank_map[!(is.na(genbank_map$refseq_mrna) | genbank_map$refseq_mrna==""),]
genbank_map <- subset(genbank_map, genbank_map$gene_biotype == "protein_coding")

# merge
master <- merge(genbank_map, annotations, by.x="refseq_mrna", by.y="Reporter.Database.Entry..genbank.")

master_sub <- master[,c(4,2)]
master_sub <- master_sub[!(is.na(master_sub$hgnc_symbol) | master_sub$hgnc_symbol==""),]

# remove multi mapping probes
master_sub <- master_sub %>% distinct(hgnc_symbol, .keep_all = T)
dim(master_sub)
```

```{r}
qnorm <- subset(ex, rownames(ex) %in% master_sub$Reporter.Name)
dim(qnorm)
```

```{r}
master_sub <- master_sub[match(rownames(qnorm), master_sub$Reporter.Name),]
rownames(qnorm) <- master_sub$hgnc_symbol
head(qnorm[1:3,1:3])
#write.csv(qnorm, "/home/clodagh/MASTERS PROJECT/GSE37614/GSE108565_qnorm", row.names =T)
qnorm <- read.csv("/home/clodagh/MASTERS PROJECT/GSE37614/GSE108565_qnorm", sep = ",", row.names = 1)
dim(qnorm)
```

```{r}
# Probe filtering
medians <- rowMedians(as.matrix(qnorm))
hist(medians, 150, col = "cornsilk1", freq = FALSE, 
             main = "Histogram of the median intensities", 
             border = "antiquewhite4", xlab = "Median intensities")
```
## Sanity check - Resistant samples should have high CD10 and GPR77

```{r}
sym <- as.data.frame(t(qnorm[c("ECE1", "C5AR2"),]))
png("/home/clodagh/MASTERS PROJECT/GSE37614/CD10 boxplot.png")
ggplot(sym, aes(x=groups, y = sym[,"ECE1"],fill = groups)) +
geom_point(size = .9, alpha = 0.7,
position = position_jitter(width = 0.2),
show.legend=F) +
  geom_boxplot(colour = "grey30",
show.legend = F,
outlier.shape = NA,
varwidth = TRUE,
size = 0.2, alpha = 0.4) +
  ggtitle("CD10+") +
# log scaled x axis to visualise low counts
scale_y_continuous(breaks = c(13,13.5,14,14,5)) +
scale_color_manual() +
  ylab("Normalized Expression") +
theme_minimal() +
  theme(
  # no need for axis title as vehicles are shown
axis.title.x = element_blank(),
axis.title.y = element_text(),
# centre and bold plot title
plot.title = element_text(hjust = 0.5, face = "bold", size = 15),
plot.margin = margin(t = 4, r = 10, b = 4, l = 4, "pt")) +
stat_compare_means(method = "t.test")
dev.off()



```


# Samples
```{r}
annotation_for_heatmap <- 
  data.frame(Tissue = meta$`neo-adjuvant chemotherapy:ch1`)

row.names(annotation_for_heatmap) <- row.names(meta)

dists <- as.matrix(dist(t(qnorm), method = "manhattan"))

rownames(dists) <- row.names(meta)
hmcol <- colorRampPalette(RColorBrewer::brewer.pal(9, "YlOrRd"))(255)
colnames(dists) <- NULL
diag(dists) <- NA

pheatmap(dists, col = (hmcol), 
         annotation_row = annotation_for_heatmap,
         legend = TRUE, 
         show_rownames = F,
         treeheight_row = 0,
         legend_breaks = c(min(dists, na.rm = TRUE), 
                         max(dists, na.rm = TRUE)), 
         legend_labels = (c("small distance", "large distance")),
         main = "Sample Clustering RMA normalised samples")
```

## pca PLot 
```{r}
PCA <- prcomp(t(log2(qnorm)), scale = TRUE)
percentVar <- round(100*PCA$sdev^2/sum(PCA$sdev^2),1)
sd_ratio <- sqrt(percentVar[2] / percentVar[1])

dataGG <- data.frame(PC1 = PCA$x[,1], PC2 = PCA$x[,2],
                    group = meta$`neo-adjuvant chemotherapy:ch1`)

ggscatter(dataGG, x="PC1", y="PC2",
                  color = "group", palette = c("dodgerblue4", "darkorange2", "chartreuse", "purple"),
                  title = "PCA plot log-transformed quantile normalized expression data",
                  subtitle = "Resistent vs. Sensitive",
                  xlab = paste0("PC1, VarExp: ", percentVar[1], "%"),
                  ylab = paste0("PC2, VarExp: ", percentVar[2], "%"),
                  ellipse = T, star.plot = T, 
                  ggtheme = theme_bw()) + 
                  theme(legend.position = "right") + 
                  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```





```{r}
lab <- paste(meta$`neo-adjuvant chemotherapy:ch1`)
colors <- as.numeric(factor(meta$`neo-adjuvant chemotherapy:ch1`))
plotMDS(qnorm, labels=lab, col=colors)
```



# Differential Expression


```{r}
design <- model.matrix(~0 + groups)
colnames(design) <- levels(gs)
aw <- arrayWeights(gset, design)  
barplot(aw)
fit <- lmFit(qnorm, design)  # fit linear model
# set up contrasts of interest and recalculate model coefficients
cts <- "resistant-sensitive"
cont.matrix <- makeContrasts(contrasts=cts, levels=design)
fit2 <- contrasts.fit(fit, cont.matrix)
```

```{r}
# compute statistics and table of top significant genes
fit2 <- eBayes(fit2, 0.01)
res <- decideTests(fit2)
```


# Differentially Expressed Genes
```{r}
tT <- topTable(fit2, adjust="fdr", sort.by="B", number=Inf)
tT <- subset(tT, select=c("logFC","adj.P.Val","P.Value","t","B","logFC"))
#write.csv(tT, file="/home/clodagh/MASTERS PROJECT/GSE37614/GSE108565_tT.csv", row.names=T)

sig <- tT[tT$adj.P.Val <= 0.05 & abs(tT$logFC)>= 1, ];
#write.table(sig, "GSE108565_DEG_list.csv", sep="\t", row.names=FALSE)
up <- tT[tT$adj.P.Val <= 0.05 & tT$logFC>= 1, ];
#write.table(up, "GSE108565_DEG_upreg_list.csv", sep="\t", row.names=FALSE)
down <- tT[tT$adj.P.Val <= 0.05 & tT$logFC <= -1, ];
#write.table(down, "GSE108565_DEG_upreg_list.csv", sep="\t", row.names=FALSE)
#Extract the expression values for the DEGs
sig_exprs <- qnorm[rownames(qnorm) %in% as.character(rownames(sig)),]
#write.table(sig_exprs, "GSE108565_DEG_counts.csv", sep="\t", row.names=TRUE)
Gene <- rownames(sig_exprs)
dim(sig_exprs)
```

# Visualise DEGs


```{r}
#Create volcano plot for DEGs.
plot(tT$logFC, -log10(tT$adj.P.Val), pch="*", xlab="Log2 Fold Change", ylab="-10log (adjusted p-value)", main = "GSE108565 DEGs")
abline(h=-log10(0.05), v=c(-1, 1), col="red", lty=2)
points(sig$logFC, -log10(sig$adj.P.Val), col="red", pch="*")
```

```{r}

#png("/home/clodagh/MASTERS PROJECT/GSE37614/GSE108565_clustering_heatmap.png")
title = "GSE108565 398 DEGs"
pheatmap(log2(sig_exprs[,ord]), trace="none", scale="row", cexRow=0.2, cexCol=0.7, annotation_col = annotation_for_heatmap, show_rownames = F, show_colnames = F, cluster_cols = T, main = title)
#dev.off()
```

# Gene Ontology

```{r}
list <- tT$adj.P.Val
names(list) <- rownames(tT)

# selection is function that returns TRUE/FALSE for p-values<0.05
selection <- function(x) TRUE

allGO2genes <- annFUN.org(whichOnto="BP", feasibleGenes=NULL, mapping="org.Hs.eg.db", ID="symbol")

GOdata <- new("topGOdata", ontology="BP", allGenes=list, annot=annFUN.GO2genes, GO2genes=allGO2genes, geneSel=selection, nodeSize=10)

```
## Perform Enrichment 
```{r}
results.ks <- runTest(GOdata, algorithm="classic", statistic="ks")
#enrichment_barplot(GOdata, result = results.ks, showTerms = 10, numChar = 40, orderBy = "Scores", y = "Count",  xlab = NULL, ylab = NULL, title = "GSE108565 GO Biological Process")
```


```{r}
goEnrichment <- GenTable(GOdata, KS=results.ks, orderBy="KS", topNodes=30)
goEnrichment$KS <- as.numeric(goEnrichment$KS)
goEnrichment <- goEnrichment[goEnrichment$KS<0.05,]
goEnrichment <- goEnrichment[,c("GO.ID","Term","KS")]
goEnrichment$Term <- gsub(" [a-z]*\\.\\.\\.$", "", goEnrichment$Term)
goEnrichment$Term <- gsub("\\.\\.\\.$", "", goEnrichment$Term)
goEnrichment$Term <- paste(goEnrichment$GO.ID, goEnrichment$Term, sep=", ")
goEnrichment$Term <- factor(goEnrichment$Term, levels=rev(goEnrichment$Term))
#write.table(goEnrichment, "/home/clodagh/MASTERS PROJECT/GSE37614/GSE108565_DEGs_GO_Enrichment", sep = "\t", row.names = F)
```

```{r}
DT::datatable(goEnrichment, options = list(scrollX = TRUE, pageLength = 10, scroller = TRUE))
```

```{r}
#Plot the GO graph, color nodes by significance
png("/home/clodagh/MASTERS PROJECT/GSE37614/GSE108565_DEGS_GOdata.png")
showSigOfNodes(GOdata, score(results.ks), firstSigNodes = 5, useInfo = "all")
```

```{r}
rm(mean)
require(ggplot2)
ggplot(goEnrichment, aes(x=Term, y=-log10(KS), fill = -log10(KS))) +
    stat_summary(geom = "bar", fun = mean, position = "dodge") +
    ylab("Enrichment (-log10(p-value))") +
    ggtitle("GSE108565 GO Biological Process") +
    scale_y_continuous(breaks = round(seq(0, max(-log10(goEnrichment$KS)), by = 2), 1)) +
    theme_bw(base_size=24) +
    theme(
        legend.position='none',
        legend.background=element_rect(),
        plot.title=element_text(angle=0, size=14, face="bold", vjust=1),
        axis.text.x=element_text(angle=0, size=8, face="bold", hjust=1.10),
        axis.text.y=element_text(angle=0, size=8, face="bold", vjust=0.5),
        axis.title.y = element_blank(),
        axis.title.x =element_text(size=14, face="bold"),
        legend.key=element_blank(),     #removes the border
        legend.key.size=unit(1, "cm"),      #Sets overall area/size of the legend
        legend.text=element_text(size=18),  #Text size
        title=element_text(size=18)) +
        guides(colour=guide_legend(override.aes=list(size=2.5))) +
        
    coord_flip()
```

```{r}
library(org.Hs.eg.db)
library(DOSE) #Disease ontology; description of gene products with disease perspectvies
library(pathview)
library(clusterProfiler)
library(AnnotationHub)
library(ensembldb)
library(tidyverse)
library(fgsea)
```
# GSEA and KEGG

```{r}
hs <- org.Hs.eg.db
#background set 
probes = rownames(qnorm)
probes2annot = AnnotationDbi::select(hs, 
       keys = probes,
       columns = c("ENTREZID", "SYMBOL"),
       keytype = "SYMBOL")
entrez <- match(probes, probes2annot$SYMBOL)
# the following is the number or probes without annotation:
sum(is.na(probes2annot))

#sig genes
sig_genes <- rownames(sig_exprs)
sig_genes <- AnnotationDbi::select(hs, 
       keys = sig_genes,
       columns = c("ENTREZID", "SYMBOL"),
       keytype = "SYMBOL")
## Run GO enrichment analysis 
ego <- enrichGO(gene = sig_genes$ENTREZID, 
                universe = probes2annot$ENTREZID,
                keyType = "ENTREZID",
                OrgDb = org.Hs.eg.db, 
                ont = "BP", #Biological Processes GO term, also done for CC and MF
                pAdjustMethod = "BH", 
                qvalueCutoff = 0.05, 
                readable = TRUE)

## Dotplot 
dotplot(ego, showCategory=10) + ggtitle("DEGs GO (Biological Process)")  +
  theme_classic()

pdf(file="/home/clodagh/MASTERSGO_enrichment_BP.pdf")
#take subset of ego_p
ego_p <- as.data.frame(ego) 
ego_p <- ego_p[1:20,]
ggplot(ego_p, aes(x = GeneRatio, y = fct_reorder(Description, GeneRatio))) + 
               geom_point(aes(size = GeneRatio, color = p.adjust)) +
               theme_bw(base_size = 14) +
        scale_colour_gradient(limits=c(0, 0.20), low="red") +
        ylab(NULL) +
        ggtitle("GO pathway enrichment")


```

# CC Degs
```{r}
## Run GO enrichment analysis 
ego <- enrichGO(gene = sig_genes$ENTREZID, 
                universe = probes2annot$ENTREZID,
                keyType = "ENTREZID",
                OrgDb = org.Hs.eg.db, 
                ont = "CC", #Biological Processes GO term, also done for CC and MF
                pAdjustMethod = "BH", 
                qvalueCutoff = 0.05, 
                readable = TRUE)

## Dotplot 
dotplot(ego, showCategory=10) + ggtitle("DEGs GO (Cellular Compartment)")  +
  theme_classic()
```




# MF Degs
```{r}
## Run GO enrichment analysis 
ego <- enrichGO(gene = sig_genes$ENTREZID, 
                universe = probes2annot$ENTREZID,
                keyType = "ENTREZID",
                OrgDb = org.Hs.eg.db, 
                ont = "MF", #Biological Processes GO term, also done for CC and MF
                pAdjustMethod = "BH", 
                qvalueCutoff = 0.05, 
                readable = TRUE)

DEGs_MF_summary <- data.frame(ego)
#write.csv(DEGs_MF_summary, "/home/clodagh/MASTERS PROJECT/GSE37614/GSE108565_DEGs_MF.csv")

## Dotplot 
dotplot(ego, showCategory=10) + ggtitle("DEGs GO (Molecular Function)")  +
  theme_classic()
```

```{r}
## To color genes by log2 fold changes, we need to extract the log2 fold changes from our results table creating a named vector
library(fgsea)
library(ggnewscale)
OE_foldchanges <-  sig$logFC
names(OE_foldchanges) <- rownames(sig)

## Cnetplot details the genes associated with one or more terms - by default gives the top 5 significant terms (by padj)
cnetplot(ego, 
         categorySize="pvalue", 
         showCategory = 5, 
         foldChange=OE_foldchanges, 
         vertex.label.font=6)
```
GSEA analysis 
Determine if a ranked gene list is significantly enriched in a prior defined biological processes/gene sets
ORA only need vector of gene IDs, GSEA needs ranked list
Create ranked list in R for use with GSEA app (ranked based on signed p-value)

```{r}
#Prepare file for GSEA input; input ALL genes for testing (including middle ranked (non-significant as removing can reduce power)
#rank genes based on signed pval
tT <- read.csv("/home/clodagh/MASTERS PROJECT/GSE37614/GSE108565_tT.csv", row.names =1)
x <- tT
x <- add_rownames(x, var = "symbol") #row to col 
head(x)
x$fcsign <- sign(x$logFC) #returns col of sign (+/- of LFC)
x$logP=-log10(x$P.Value) #take log of all p val
x$metric= x$logP/x$fcsign #signed pval ranking
x <- x[order(-x$metric),] #order in descending order
x <-  na.omit(x) #remove rows with na 
dim(x)
x <- x[!duplicated(x$symbol),] #remove duplicates
dim(x)

res <- x
```



```{r}
#write the lists - used both as loss of some genes when mapping to gene symbols. No row or col names for .rnk file
write.table(x,file="/home/clodagh/MASTERS PROJECT/GSE37614/GSE108565_de_genes_entrezid.rnk",quote=F,sep="\t",row.names=F, col.names = F)

```

Code for plotting GSEA analysis results 
Read in results of GSEA analysis and plot significant results (p.adj < 0.05

## REfine BIO tutorial
```{r}
# Attach the library
library(clusterProfiler)

# Package that contains MSigDB gene sets in tidy format
library(msigdbr)

# Zebrafish annotation package we'll use for gene identifier conversion
library(org.Hs.eg.db)

# We will need this so we can use the pipe: %>%
library(magrittr)
```




```{r}
# First let's create a mapped data frame we can join to the differential
# expression stats
dge_mapped_df <- data.frame(
  entrez_id = mapIds(
    # Replace with annotation package for the organism relevant to your data
    org.Hs.eg.db,
    keys = x$symbol,
    # Replace with the type of gene identifiers in your data
    keytype = "SYMBOL",
    # Replace with the type of gene identifiers you would like to map to
    column = "ENTREZID",
    # This will keep only the first mapped value for each symbol
    multiVals = "first"
  )
) %>%
  # If a symbol doesn't map to a Entrez gene identifier,
  # drop that from the data frame
  dplyr::filter(!is.na(entrez_id)) %>%
  # Make an `Ensembl` column to store the rownames
  tibble::rownames_to_column("SYMBOL") %>%
  # Now let's join the rest of the expression data
  dplyr::inner_join(x, by = c("SYMBOL" = "symbol"))

dge_mapped_df

#write.csv(dge_mapped_df, "/home/clodagh/MASTERS PROJECT/GSE37614/GSE108565_dge_mapped_GSEA_table.csv")
```


```{r}
any(duplicated(dge_mapped_df$entrez_id))
```

```{r}
# Let's create a named vector ranked based on the t-statistic values
t_vector <-dge_mapped_df$t
names(t_vector) <- dge_mapped_df$entrez_id

# We need to sort the t-statistic values in descending order here
t_vector <- sort(t_vector, decreasing = TRUE)

# Look at first entries of the ranked t-statistic vector
head(t_vector)
```
# Run GSEA
```{r}
msigdbr_collections()

dr_hallmark_df <- msigdbr(
  species = "Homo sapiens", # H for hallmark
  category = "H"
)


gsea_results <- GSEA(
  geneList = t_vector, # Ordered ranked gene list
  minGSSize = 25, # Minimum gene set size
  maxGSSize = 500, # Maximum gene set set
  pvalueCutoff = 0.05, # p-value cutoff
  eps = 0, # Boundary for calculating the p-value
  seed = TRUE, # Set seed to make results reproducible
  pAdjustMethod = "BH", # Benjamini-Hochberg correction
  TERM2GENE = dplyr::select(
    dr_hallmark_df,
    gs_name,
    entrez_gene
  )
)
```

```{r}
# We can access the results from our gseaResult object using `@result`
head(gsea_results@result)
gsea_result_df <- data.frame(gsea_results@result)
filter_df <- gsea_result_df[, c("ID", "enrichmentScore", "NES", "p.adjust", "qvalues")]
#write.csv(filter_df, "/home/clodagh/MASTERS PROJECT/GSE37614/DEGs_enriched_hallmarks.csv", row.names=F)
# Visualize 
#Most Positive NES
gsea_result_df %>%
  # This returns the 3 rows with the largest NES values
  dplyr::slice_max(n = 3, order_by = NES)

most_positive_nes_plot <- enrichplot::gseaplot(
  gsea_results,
  geneSetID = "HALLMARK_INTERFERON_GAMMA_RESPONSE",
  title = "HALLMARK_INTERFERON_GAMMA_RESPONSE",
  color.line = "#0d76ff"
)

most_positive_nes_plot
```

```{r}
ggplot2::ggsave(file.path("/home/clodagh/MASTERS PROJECT/GSE37614/GSE108565_gsea_enrich_positive_plot.png"),
  plot = most_positive_nes_plot
)
```

# Most negative 
```{r}
gsea_result_df %>%
  # Return the 3 rows with the smallest (most negative) NES values
  dplyr::slice_min(n = 3, order_by = NES)

most_negative_nes_plot <- enrichplot::gseaplot(
  gsea_results,
  geneSetID = "HALLMARK_MYOGENESIS",
  title = "HALLMARK_MYOGENESIS",
  color.line = "#0d76ff"
)

most_negative_nes_plot
```


```{r}
readr::write_tsv(
  gsea_result_df,
  file.path(
    "/home/clodagh/MASTERS PROJECT/GSE37614/GSE108565gsea_results.tsv"
  )
)
```


# Plot significant results
```{r}
pdf(file="/home/clodagh/MASTERS PROJECT/GSE37614/GSE108565_sig_hallmark_pathways.pdf")
#too many terms so just plot significant ones
hallmark <-gsea_result_df[gsea_result_df$p.adjust<0.05,]
ggplot(hallmark, aes(reorder(ID, NES), NES)) +
  geom_col(aes(fill=NES < 0)) +
  #scale_fill_manual(name = '', values = setNames(c("#F46D43","#74ADD1"),c("", "")), guide = NULL) +
  scale_fill_manual(values = c("#fa624b","#5e61fc")) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
       title="Hallmark pathways NES from GSEA") + 
  theme_classic() 
dev.off()
```

# Run GSEA
```{r}
msigdbr_collections()

dr_hallmark_df <- msigdbr(
  species = "Homo sapiens", # H for hallmark
  category = "C2", #C5 for GO
  subcategory = "CP:KEGG"
)


gsea_results <- GSEA(
  geneList = t_vector, # Ordered ranked gene list
  minGSSize = 25, # Minimum gene set size
  maxGSSize = 500, # Maximum gene set set
  pvalueCutoff = 0.05, # p-value cutoff
  eps = 0, # Boundary for calculating the p-value
  seed = TRUE, # Set seed to make results reproducible
  pAdjustMethod = "BH", # Benjamini-Hochberg correction
  TERM2GENE = dplyr::select(
    dr_hallmark_df,
    gs_name,
    entrez_gene
  )
)

gsea_result_df <- data.frame(gsea_results@result)
#write.csv(gsea_result_df, "/home/clodagh/MASTERS PROJECT/GSE37614/GSE108565_KEGG_enrichment DEGs")

png("/home/clodagh/MASTERS PROJECT/GSE37614/GSE108565_KEGG_barplotDEGs.png")
KEGG <-gsea_result_df[gsea_result_df$p.adjust<0.0005,]
ggplot(KEGG, aes(reorder(ID, NES), NES)) +
  geom_col(aes(fill=NES < 0)) +
  #scale_fill_manual(name = '', values = setNames(c("#F46D43","#74ADD1"),c("", "")), guide = NULL) +
  scale_fill_manual(values = c("#fa624b","#5e61fc")) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
       title="KEGG NES from GSEA") + 
  theme_classic() 
dev.off()

```







# KEGG
```{r}
gsea_results <- gseKEGG(
  geneList = t_vector, # Ordered ranked gene list
  organism = "hsa",
  keyType = "kegg",
  minGSSize = 25, # Minimum gene set size
  maxGSSize = 500, # Maximum gene set set
  pvalueCutoff = 0.05, # p-value cutoff
  eps = 0, # Boundary for calculating the p-value
  seed = TRUE, # Set seed to make results reproducible
  pAdjustMethod = "BH", # Benjamini-Hochberg correction
  by = "fgsea"
  )
gsea_result_KEGG <- data.frame(gsea_results@result)
gsea_result_KEGG <- gsea_result_KEGG %>% arrange(-NES)
keggp <- gsea_result_KEGG[gsea_result_KEGG$p.adjust<0.000000005,]
#write.csv(keggp, "/home/clodagh/MASTERS PROJECT/GSE37614/GSE108565_KEGG_pathways.csv")
ggplot(keggp, aes(reorder(Description, NES), NES)) +
  geom_col(aes(fill=NES < 0)) +
  #scale_fill_manual(name = '', values = setNames(c("#F46D43","#74ADD1"),c("", "")), guide = NULL) +
 scale_fill_manual(values = c("#fa624b","#5e61fc")) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
       title="KEGG NES from GSEA") + 
  theme_classic() 
dev.off()
```



