---
title: "GSE37614 batch correction"
author: "Clodagh Murray"
date: "5/19/2022"
output: html_document
---


### Libraries
```{r, message =F}
library(oligo)
library(Biobase)
library(ggpubr)
library(data.table)
library(DT)
library(dplyr)
library(ggpubr)
library(ggplot2)
library(limma)
library(RColorBrewer)
library(pheatmap)
library(PCAtools)
library(illuminaHumanv4.db)
```

## Raw Data Staging
### 4 samples were analysed in both batches 1 and 2, creating 4 extra samples in the raw data - update metadata to match raw data 

```{r}
SDRF <- read.delim("/home/clodagh/MASTERS PROJECT/GSE37614/E-GEOD-37614.sdrf_fixed.txt")
rownames(SDRF) <- SDRF$Array.Data.File
DT::datatable(SDRF, options = list(scrollX = TRUE, pageLength = 23, scroller = TRUE))
```

## Count matrix matching metadata?
```{r}
## same as first attempt tbh
gen_raw <- function(path, pattern){
  files = list.files(path, pattern, full.names=T, recursive=T, include.dirs=T)
  mat = as.data.frame(do.call(cbind, lapply(files, function(x) data.table::fread(x, stringsAsFactors=FALSE))))
  ID_REF = mat[,1]
  mat = as.data.frame(mat[,!grepl("ID_REF", colnames(mat))])
  rownames(mat) = ID_REF
  return(mat)
}

raw.dat <- gen_raw("/home/clodagh/MASTERS PROJECT/GSE37614/raw/", "\\.txt$")

## attach sample names to cols
file_names <- list.files("/home/clodagh/MASTERS PROJECT/GSE37614/raw/", "\\.txt$", full.names=T, recursive=T, include.dirs=T)
file_names <- sub(".*/", "", file_names)
sample_names <- sub("\\_.*", "", file_names)
length(sample_names) == ncol(raw.dat)
```

## Append IDs to matrix

```{r}
#  append batch ID's to the sample names for unique colnames. 
SDRF$samp_names <- paste(SDRF$Assay.Name, SDRF$Characteristics..batch., sep=".")

```
```{r}
# make sure columns match file contents
SDRF <- SDRF[match(file_names, SDRF$Array.Data.File),]
colnames(raw.dat) <- SDRF$samp_names
samples <- SDRF

## the below is correct 
head(raw.dat[1:5, 1:5])
```

## Make Expression Set manually

```{r}
SDRF <- Biobase::AnnotatedDataFrame(SDRF)
rownames(SDRF) <- colnames(raw.dat)
my_set <- Biobase::ExpressionSet(assayData = as.matrix(raw.dat), phenoData = SDRF)
```

## Raw Data Boxplot
```{r}
SDRF@data$FactorValue..PATIENT.ID.
SDRF@data$Characteristics..class.[23] <- "Her2+"
sml <- as.character(as.numeric(as.factor(SDRF@data$Characteristics..class.))) 
gs <- factor(sml)
groups <- SDRF@data$Characteristics..class. 
orientation <- unique(groups)
orientation <- c(orientation[2], orientation[3], orientation[1])
levels(gs) <- orientation
my_set@phenoData$group <- gs
ord <- order(gs)

palette <- palette(c("#1B9E77", "#7570B3", "pink"))

par(mar=c(7,4,2,1))
title <- "GSE37614 Log2 Raw Signals"
boxplot(log2(exprs(my_set[,ord])), boxwex=0.6, notch=T, main=title, outline=FALSE, las=2, col=gs[ord])
legend("topleft", orientation, fill=palette(), bty="n")
```


## Quantile normalization

```{r}
quantile_normalisation <- function(df){
  df_rank <- apply(df,2,rank,ties.method="min")
  df_sorted <- data.frame(apply(df, 2, sort))
  df_mean <- apply(df_sorted, 1, mean)

  index_to_mean <- function(my_index, my_mean){
    return(my_mean[my_index])
  }
df_final <- apply(df_rank, 2, index_to_mean, my_mean=df_mean)
  rownames(df_final) <- rownames(df)
  return(df_final)
}

qnorm <- quantile_normalisation(raw.dat)
```


## Update ExpressionSet Object 

### Boxplot of Qnorm

```{r}
exprs(my_set) <- qnorm

## boxplot
par(mar=c(7,4,2,1))
title <- "GSE37614 Log2 Quantile Normalization"
boxplot(log2(qnorm[,ord]), boxwex=0.6, notch=T, main=title, outline=FALSE, las=2, col=gs[ord])
legend("topleft", orientation, fill=palette(), bty="n")
```
## Expression Distribution

## Probe Filtering

```{r}
# Probe filtering
medians <- rowMedians(ex)

hist(log2(medians + 1), 150, col = "cornsilk1", freq = FALSE, 
             main = "Histogram of the median intensities", 
             border = "antiquewhite4", xlab = "Median intensities")
```


## Remove Multimapping probes

```{r, message = F}
probes <- rownames(my_set)
```

```{r}
annotation <- AnnotationDbi::select(illuminaHumanv4.db,
                                    keys = probes,
                                    columns = c("SYMBOL", "GENENAME", "ENTREZID"),
                                    keytype = "PROBEID")
```

```{r}
## 'select()' returned 1:many mapping between keys and columns
annotation <- subset(annotation, !is.na(ENTREZID))

## resolve multi maps. 

ann_grouped <- group_by(annotation, PROBEID)
ann_sum <- dplyr::summarize(ann_grouped, no_of_matches = n_distinct(ENTREZID))
ann_flt <- dplyr::filter(ann_sum, no_of_matches > 1)
remove_id <- (probes %in% ann_flt$PROBEID)

table(remove_id) # still too many probes
```
### Too many genes
*manual filtering*

```{r}
annotations <- read.table("/home/clodagh/MASTERS PROJECT/GSE37614/A-GEOD-10558.adf.txt", header=T, sep="\t")
mart <- biomaRt::useMart(biomart = "ensembl", dataset = "hsapiens_gene_ensembl")
attr <- biomaRt::listAttributes(mart)
genbank_map <- biomaRt::getBM(attributes = c("refseq_mrna", "hgnc_symbol", "gene_biotype"), mart = mart, useCache = FALSE)
genbank_map <- genbank_map[!(is.na(genbank_map$refseq_mrna) | genbank_map$refseq_mrna==""),]
genbank_map <- subset(genbank_map, genbank_map$gene_biotype == "protein_coding")

# pop off the version ID
annotations$genbank <- annotations$Reporter.Database.Entry..genbank.

annotations$genbank <- gsub("\\.\\d+$", "", annotations$genbank)
# merge
master <- merge(genbank_map, annotations, by.x="refseq_mrna", by.y="genbank")
```

```{r}
master_sub <- master[,c(4,2)]
master_sub <- master_sub[!(is.na(master_sub$hgnc_symbol) | master_sub$hgnc_symbol==""),]

# remove multi mapping probes
master_sub <- master_sub %>% distinct(hgnc_symbol, .keep_all = T)
dim(master_sub)

```

```{r}
qnorm <- subset(qnorm, rownames(qnorm) %in% master_sub$Reporter.Name)
dim(qnorm)
```

```{r}
master_sub <- master_sub[match(rownames(qnorm), master_sub$Reporter.Name),]
rownames(qnorm) <- master_sub$hgnc_symbol

# qnorm now annotated with gene symbol instead of reporter name
head(qnorm[1:3,1:3])
```



```{r}
medians <- rowMedians(qnorm)

hist(log2(medians + 1), 150, col = "cornsilk1", freq = FALSE, 
             main = "Histogram of the median intensities", 
             border = "antiquewhite4", xlab = "Median intensities")
```
# Cant have bimodal distribution
```{r}
medians <- log2(rowMedians(qnorm) + 1)
qnorm <- as.data.frame(qnorm)
qnorm$medians <- medians

qnorm <- subset(qnorm, qnorm$medians > 7.9)
qnorm <- as.matrix(qnorm[,1:(ncol(qnorm)-1)]) # get rid of last col (medians) 

# plot again
hist(log2(qnorm + 1), 150, col = "cornsilk1", freq = FALSE, 
             main = "Histogram of the median intensities", 
             border = "antiquewhite4", xlab = "Median intensities")
dim(qnorm)
#write.csv(qnorm, "/home/clodagh/MASTERS PROJECT/GSE37614/GSE37614_qnorm.csv", row.names = T)
#qnorm <- read.csv("/home/clodagh/MASTERS PROJECT/GSE37614/GSE37614_qnorm.csv", row.names = 1)

```
# Exploratory Data Analysis
## Sample heatmap

```{r}
meta <- my_set@phenoData@data
meta <- meta[,c(4, 5)]
colnames(meta) <- c("Batch", "Cancer_Subtype")
phenotype_names <- as.factor(meta$Cancer_Subtype)
batch <- meta$Batch
annotation_for_heatmap <- 
  data.frame(subtype = phenotype_names,  Batch= batch)
row.names(annotation_for_heatmap) <- rownames(meta)
```



```{r}
dists <- as.matrix(dist(t(qnorm), method = "manhattan"))

rownames(dists) <- rownames(meta)
hmcol <- colorRampPalette(RColorBrewer::brewer.pal(9, "YlOrRd"))(255)
colnames(dists) <- NULL
diag(dists) <- NA

ann_colors <- list(
  subtype= c("TNBC" = "chartreuse4", "Her2+" = "burlywood3", "ER+" = "purple"),
  Batch = c(b1= "blue4", b2 = "cadetblue2")
                   )
pheatmap(dists, col = hmcol, cluster_rows= T,
         annotation_row = annotation_for_heatmap,
         annotation_colors = ann_colors,
         legend = TRUE, 
         treeheight_row = 0,
         legend_breaks = c(min(dists, na.rm = TRUE), 
                         max(dists, na.rm = TRUE)), 
         legend_labels = (c("small distance", "large distance")),
         main = "Clustering heatmap for the calibrated samples")
```

```{r}
qnorm <- qnorm[,-23]
meta <- my_set@phenoData@data
meta <- meta[-23,] #remove outlier
meta <- meta[,c(4, 5)]
colnames(meta) <- c("Batch", "Cancer_Subtype")
#write.table(meta, "metadata.csv", row.names=T, sep="\t")
#write.csv(qnorm, "/home/clodagh/MASTERS PROJECT/GSE37614/GSE37614_qnorm.csv", row.names=T)
```

-----------------------------------------------------------------------
#### BEGIN HERE #############
```{r}
qnorm <- read.csv("/home/clodagh/MASTERS PROJECT/GSE37614/GSE37614_qnorm.csv", row.names=1)
meta <-  read.csv("/home/clodagh/MASTERS PROJECT/GSE37614/metadata.csv", sep = "\t")
```


# Exploratory Data Analysis
## Sample heatmap

```{r}
phenotype_names <- meta$Cancer_Subtype
batch <- meta$Batch
annotation_for_heatmap <- 
  data.frame(subtype = phenotype_names,  Batch= batch)
row.names(annotation_for_heatmap) <- rownames(meta)
```



```{r}
dists <- as.matrix(dist(t(qnorm), method = "manhattan"))

rownames(dists) <- rownames(meta)
hmcol <- colorRampPalette(RColorBrewer::brewer.pal(9, "YlOrRd"))(255)
colnames(dists) <- NULL
diag(dists) <- NA

ann_colors <- list(
  subtype= c("TNBC" = "chartreuse4", "Her2+" = "burlywood3", "ER+" = "purple"),
  Batch = c(b1= "blue4", b2 = "cadetblue2")
                   )
pheatmap(dists, col = (hmcol), cluster_rows= T,
         annotation_row = annotation_for_heatmap,
         annotation_colors = ann_colors,
         legend = TRUE, 
         treeheight_row = 0,
         legend_breaks = c(min(dists, na.rm = TRUE), 
                         max(dists, na.rm = TRUE)), 
         legend_labels = (c("small distance", "large distance")),
         main = "Clustering heatmap for the calibrated samples")
```



### PCA Cancer Subtypes
*2 groups - largely driven by batch*

```{r}
PCA <- prcomp(t(log2(qnorm)), scale = TRUE)
percentVar <- round(100*PCA$sdev^2/sum(PCA$sdev^2),1)
sd_ratio <- sqrt(percentVar[2] / percentVar[1])

dataGG <- data.frame(PC1 = PCA$x[,1], PC2 = PCA$x[,2],
                    Subtype = meta$Cancer_Subtype,
                    Batch = meta$Batch)


ggscatter(dataGG, x="PC1", y="PC2",
                  color = "Subtype", palette = c("dodgerblue4", "darkorange2", "chartreuse", "purple"),
                  title = "PCA plot log-transformed quantile normalized expression data",
                  subtitle = "Cancer Subtypes",
                  xlab = paste0("PC1, VarExp: ", percentVar[1], "%"),
                  ylab = paste0("PC2, VarExp: ", percentVar[2], "%"),
                  ellipse = FALSE, star.plot = F,
                  ggtheme = theme_bw()) + 
                  theme(legend.position = "right") + 
                  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

### PCA Batch Effects
```{r}
dataGG <- data.frame(PC1 = PCA$x[,1], PC2 = PCA$x[,2],
                    Batch = meta$Batch)


ggscatter(dataGG, x="PC1", y="PC2",
                  color = "Batch", palette = c("dodgerblue4", "darkorange2", "chartreuse", "purple"),
                  title = "PCA plot log-transformed quantile normalized expression data",
                  subtitle = "Batch Effects (Batch 1, Batch 2)",
                  xlab = paste0("PC1, VarExp: ", percentVar[1], "%"),
                  ylab = paste0("PC2, VarExp: ", percentVar[2], "%"),
                  ellipse = TRUE, star.plot = TRUE,
                  ggtheme = theme_bw()) + 
                  theme(legend.position = "right") + 
                  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```


## Pearsons Correlation btween PCs and batch and cancer

```{r}
pca <- PCAtools::pca(log2(qnorm + 1), metadata = meta)

PCAtools::eigencorplot(pca, components = PCAtools::getComponents(pca, 1:10),
    metavars = c('Batch','Cancer_Subtype'),
    col = c('white', 'cornsilk1', 'gold', 'forestgreen', 'darkgreen'),
    cexCorval = 1.2, fontCorval = 2, posLab = 'all', rotLabX = 45,
    scale = TRUE,
    main = bquote(PC ~ Pearson ~ r^2 ~ clinical ~ correlates),
    plotRsquared = TRUE,
    corFUN = 'pearson',
    corUSE = 'pairwise.complete.obs',
    corMultipleTestCorrection = 'BH',
    signifSymbols = c('****', '***', '**', '*', ''),
    signifCutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1))
```


## Remove Batch

```{r}
groups <- as.factor(meta$Cancer_Subtype)
batch <- as.factor(meta$Batch)
design <- model.matrix(~0 + groups + batch)
gps <- c("ER", "HER", "TNBC", "b2")
colnames(design) <- gps
rm_batch <- limma::removeBatchEffect(log2(qnorm + 1), batch = meta$Batch, design=design)

#write.table(rm_batch, file = '/home/clodagh/MASTERS PROJECT/GSE37614/GSE37614_rm_batch.txt', sep ='\t', col.names = T, row.names = T)

```


## PCA Cancer Subtype (Corrected)

```{r}
PCA <- prcomp(t(log2(qnorm)), scale = TRUE)
percentVar <- round(100*PCA$sdev^2/sum(PCA$sdev^2),1)
sd_ratio <- sqrt(percentVar[2] / percentVar[1])

dataGG <- data.frame(PC1 = PCA$x[,1], PC2 = PCA$x[,2],
                    Subtype = meta$Cancer_Subtype)


ggscatter(dataGG, x="PC1", y="PC2",
                  color = "Subtype", palette = c("dodgerblue4", "darkorange2", "chartreuse", "purple"),
                  title = "PCA plot log-transformed batch corrected data",
                  subtitle = "Cancer Subtypes",
                  xlab = paste0("PC1, VarExp: ", percentVar[1], "%"),
                  ylab = paste0("PC2, VarExp: ", percentVar[2], "%"),
                  ellipse = F, star.plot = T, 
                  ggtheme = theme_bw()) + 
                  theme(legend.position = "right") + 
                  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

## PCA Batch Effect (Corrected)

```{r}
dataGG <- data.frame(PC1 = PCA$x[,1], PC2 = PCA$x[,2],
                    Batch = meta$Batch)


ggscatter(dataGG, x="PC1", y="PC2",
                  color = "Batch", palette = c("dodgerblue4", "darkorange2", "chartreuse", "purple"),
                  title = "PCA plot log-transformed batch corrected data",
                  subtitle = "Batch Effects: Batch1, Batch2",
                  xlab = paste0("PC1, VarExp: ", percentVar[1], "%"),
                  ylab = paste0("PC2, VarExp: ", percentVar[2], "%"),
                  ellipse = T, star.plot = T, 
                  ggtheme = theme_bw()) + 
                  theme(legend.position = "right") + 
                  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

## Sample heatmap (batch correcetd)


```{r}
dists <- as.matrix(dist(t(log2(qnorm)), method = "manhattan"))

rownames(dists) <- rownames(meta)
hmcol <- colorRampPalette(RColorBrewer::brewer.pal(9, "YlOrRd"))(255)
colnames(dists) <- NULL
diag(dists) <- NA

ann_colors <- list(
  subtype= c("TNBC" = "chartreuse4", "Her2+" = "burlywood3", "ER+" = "purple"),
  Batch = c(b1= "blue4", b2 = "cadetblue2"))

# Remove outlier? 
pheatmap(dists, col = (hmcol),
         annotation_row = annotation_for_heatmap,
         annotation_colors = ann_colors,
         legend = TRUE,
         treeheight_row = 0,
         legend_breaks = c(min(dists, na.rm = TRUE), 
                         max(dists, na.rm = TRUE)), 
         legend_labels = (c("small distance", "large distance")),
         main = "GSE37614 Clustering heatmap normalized samples")
```

