---
title: "microarray analysis"
output:
  html_document:
    df_print: paged
---




```{r}
library(illuminaHumanv4.db)
library(ggpubr)
library(GEOquery)
library(limma)
library(gplots)
library(dplyr)
```

## Importing the Data

```{r}
library(GEOquery)
my_id <- "GSE37614"
gse <- getGEO(my_id, GSEMatrix =TRUE)
if (length(gse) > 1) idx <- grep("GPL10558", attr(gse, "names")) else idx <- 1
gset <- gse[[idx]]
```

```{r}
pData(gset)
fData(gset)
fvarLabels(gset)
fvarLabels(gset) <- make.names(fvarLabels(gset))
features <- fData(gset)
```



```{r}
gsms <- "2221111100002221100"
sml <- strsplit(gsms, split="")[[1]]
```


## Check normalisation and scales used
Log2 scale transformation required

Check for normalisation by looking at boxplot of expression levels for each sample

```{r}
ex <- exprs(gset)
qx <- as.numeric(quantile(ex, c(0., 0.25, 0.5, 0.75, 0.99, 1.0), na.rm=T))
LogC <- (qx[5] > 100) ||
  (qx[6]-qx[1] > 50 && qx[2] > 0)
if (LogC) { ex[which(ex <= 0)] <- NaN
exprs(gset) <- log2(ex) }


gs <- factor(sml)
groups <- make.names(c("HER","ER","TNBC"))
levels(gs) <- groups
gset$group <- gs
ord <- order(gs)  # order samples by group

palette <- palette(c("#1B9E77", "#7570B3", "#E7298A", "#E6AB02", "#D95F02",
          "#66A61E", "#A6761D", "#B32424", "#B324B3", "#666666"))
par(mar=c(7,4,2,1))
title <- paste ("GSE37614", "/", annotation(gset), sep ="")
boxplot(log2(ex[,ord]), boxwex=0.6, notch=T, main=title, outline=FALSE, las=2, col=gs[ord])
legend("topleft", groups, fill=palette(), bty="n")
```


## PCA plot



library(ggpubr)

ex <- exprs(gset)
qx <- as.numeric(quantile(ex, c(0., 0.25, 0.5, 0.75, 0.99, 1.0), na.rm=T))
LogC <- (qx[5] > 100) ||
  (qx[6]-qx[1] > 50 && qx[2] > 0)
if (LogC) { ex[which(ex <= 0)] <- NaN
exprs(gset) <- log2(ex) }



PCA <- prcomp(t(expression_set), scale = FALSE)

percentVar <- round(100*PCA$sdev^2/sum(PCA$sdev^2),1)
sd_ratio <- sqrt(percentVar[2] / percentVar[1])

dataGG <- data.frame(PC1 = PCA$x[,1], PC2 = PCA$x[,2],
                    Subtype = gset$group)

ggpubr::ggscatter(dataGG, x="PC1", y="PC2",
                  color = "Subtype", palette = c("dodgerblue4", "darkorange2", "cyan3"),
                  title = "PCA plot log-transformed RMA normalized expression data",
                  xlab = paste0("PC1, VarExp: ", percentVar[1], "%"),
                  ylab = paste0("PC2, VarExp: ", percentVar[2], "%"),
                  ellipse = TRUE, star.plot = TRUE,
                  ggtheme = theme_bw()) + 
                  theme(legend.position = "right") + 
                  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

## Probe filtering

```{r}

# expression value distribution
par(mar=c(4,4,2,1))
title <- paste ("Normalized Expression Density", sep ="")
plotDensities(ex[,ord], group=gs, main=title, legend ="topright")

# Probe filtering
medians <- rowMedians(ex)
man_threshold <- 4

hist(medians, 150, col = "cornsilk1", freq = FALSE, 
main = "Histogram of the median intensities", 
border = "antiquewhite4",
xlab = "Median intensities")

```

```{r}
meta <- pData(gset)
meta$group
no_of_samples <- table(meta$group)
sample_cutoff <- min(no_of_samples)

idx_man_threshold <- apply(Biobase::exprs(gset), 1, function(x){ sum(x > man_threshold) >= sample_cutoff})

table(idx_man_threshold)

probes <- rownames(gset)

library(illuminaHumanv4.db)
annotation <- AnnotationDbi::select(illuminaHumanv4.db,
                                    keys = probes,
                                    columns = c("SYMBOL", "GENENAME"),
                                    keytype = "PROBEID")
annotation <- subset(annotation, !is.na(SYMBOL))

## resolve multi maps. 

ann_grouped <- group_by(annotation, PROBEID)
ann_sum <- dplyr::summarize(ann_grouped, no_of_matches = n_distinct(SYMBOL))

ann_flt <- filter(ann_sum, no_of_matches > 1)

remove_id <- (featureNames(gset) %in% ann_flt$PROBEID)

table(remove_id)
print("542 PROBES map to multiple gene IDs %% REMOVING PROBES")
```

```{r}
final_annotation <- subset(annotation, !remove_id)
final_Obj <- subset(gset, !remove_id)

fData(final_Obj)$PROBEID <- rownames(fData(final_Obj))

fData(final_Obj) <- left_join(fData(final_Obj), annotation)

rownames(fData(final_Obj)) <- fData(final_Obj)$PROBEID
```

# Design matrix
```{r}
design <- model.matrix(~group + 0, gset)
colnames(design) <- levels(gs)
fit <- lmFit(gset, design)

# Set up contrasts of interest
cts <- paste(groups, c(tail(groups, -1), head(groups, 1)), sep="-")
cont.matrix <- makeContrasts(contrasts=cts, levels=design)
fit2 <- contrasts.fit(fit, cont.matrix)
fit2 <- eBayes(fit2, 0.01)

#Create the top table for the comparison (coef) we want
top <- topTable(fit2, coef=1, adjust="fdr", sort.by="B", number = Inf)
nrow(top)

#Top Table of significant genes only
up_reg_epi <- subset(top, logFC > 0.25 & P.Value < 0.05)
nrow(up_reg_epi)

#Create an excel-readable table with the specified columns for sig gene only
tT <- subset(up_reg_epi, select=c("ID","adj.P.Val","P.Value","t","Symbol","ILMN_Gene"))
write.table(tT, "DEG_list.csv", row.names=F, sep="\t")

#Sanity check - hist of all P vals
hist(top$adj.P.Val, col = "grey", border = "white", xlab = "P-adj",
     ylab = "Number of genes", main = "P-adj value distribution")

# Hist of sig DEG P values
hist(up_reg_epi$adj.P.Val, col = "grey", border = "white", xlab = "P-adj",
     ylab = "Number of genes", main = "P-adj value distribution")

#Extract the expression values for the DEGs
e <- exprs(gset)
sig_exprs <- e[rownames(e) %in% as.character(tT$ID),]
```

# Up-Regulated Genes
```{r}
library(DT)
DT::datatable(up_reg_epi, rownames = FALSE, options=list(scrollX=T))
```



```{r}
# summarize test results as "up", "down" or "not expressed"
dT <- decideTests(fit2, adjust.method="fdr", p.value=0.05)
vennDiagram(dT, circle.col=palette())
```

```{r}
# create Q-Q plot for t-statistic
t.good <- which(!is.na(fit2$F)) # filter out bad probes
qqt(fit2$t[t.good], fit2$df.total[t.good], main="Moderated t statistic")
```

```{r}
# volcano plot (log P-value vs log fold change)
colnames(fit2) # list contrast names
ct <- 1
# choose contrast of interest
volcanoplot(fit2, coef=ct, main=colnames(fit2)[ct], pch=20,
            highlight=length(which(dT[,ct]!=0)), names=rep('+', nrow(fit2)))

ct <- 2
# choose contrast of interest
volcanoplot(fit2, coef=ct, main=colnames(fit2)[ct], pch=20,
            highlight=length(which(dT[,ct]!=0)), names=rep('+', nrow(fit2)))

ct <- 3
# choose contrast of interest
volcanoplot(fit2, coef=ct, main=colnames(fit2)[ct], pch=20,
            highlight=length(which(dT[,ct]!=0)), names=rep('+', nrow(fit2)))


```


# MD plot (log fold change vs mean log expression)
# highlight statistically significant (p-adj < 0.05) probes
```{r}
ct <- 1
plotMD(fit2, column=ct, status=dT[,ct], legend=F, pch=20, cex=1)
abline(h=0)

ct <- 2
plotMD(fit2, column=ct, status=dT[,ct], legend=F, pch=20, cex=1)
abline(h=0)

ct <- 3
plotMD(fit2, column=ct, status=dT[,ct], legend=F, pch=20, cex=1)
abline(h=0)
```


# Heat Map

```{r}
library(pheatmap)
annotation_for_heatmap <- 
  data.frame(Subtype = gset$group)

dists <- as.matrix(dist(t(exprs(gset)), method = "manhattan"))
rownames(dists) <- row.names(annotation_for_heatmap)
hmcol <- rev(colorRampPalette(RColorBrewer::brewer.pal(9, "YlOrRd"))(255))
colnames(dists) <- NULL
diag(dists) <- NA

ann_colors <- list(
  Subtype = c(HER = "purple", ER = "forestgreen", TNBC = "cyan3"))

pheatmap(dists, col = (hmcol), 
         annotation_row = annotation_for_heatmap,
         annotation_colors = ann_colors,
         legend = TRUE, 
         show_rownames = FALSE,
         treeheight_row = 0,
         legend_breaks = c(min(dists, na.rm = TRUE), 
                         max(dists, na.rm = TRUE)), 
         legend_labels = (c("small distance", "large distance")),
         main = "Clustering heatmap RMA normalised samples")
```





